A new tidy data structure to support exploration and modeling of temporal data

### 2019-05-15 main.Rmd
remove \autoref

---
title: "A new tidy data structure to support exploration and modeling of temporal data"
author:
  - familyname: Wang
    othernames: Earo
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : earo.wang@monash.edu
    correspondingauthor: true
  - familyname: Cook
    othernames: Dianne
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : dicook@monash.edu
  - familyname: Hyndman
    othernames: Rob J
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : rob.hyndman@monash.edu
abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, and heterogeneous data types. On the other hand, the software supporting time series modeling and forecasting, makes strict assumptions on the data to be provided, typically requiring a matrix of numeric data with implicit time indexes. Going from raw data to model-ready data is painful. This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization, modeling and forecasting routines. Tidy data principles are extended to temporal data by: (1) mapping the semantics of a dataset into its physical layout; (2) including an explicitly declared index variable representing time; (3) incorporating a "key" comprising single or multiple variables to uniquely identify units over time. This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based contexts. A sound data pipeline facilitates a fluent workflow for analyzing temporal data. The infrastructure of tidy temporal data has been implemented in the R package **tsibble**.'
keywords: "time series, data wrangling, tidy data, R, forecasting, data science, exploratory data analysis, data pipelines"
wpnumber: no/yr
blind: false
cover: false
toc: false
bibliography: [references.bib]
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
library(knitr)
opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE,
  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center',
  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
  fig.height = 5, fig.width = 8, out.width = "\\textwidth",
  comment = "#>"
)
read_chunk('scripts/demo.R')
read_chunk('scripts/flights.R')
read_chunk('scripts/smart-meter.R')
```

@@

<!--Suppose that temporal data consists of $N$ observational units. Given observation $i$ $(i = 1, 2, \dots, N)$, $X_{ijt}$ denotes the $j$th measurement $(j = 1, 2, \dots, p)$ indexed at a unique sequence of time points $0 \le t \le T_{i}$.
-->
Temporal data arrives in many possible formats, with many different time contexts. For example, time can have various resolutions (hours, minutes, and seconds), and can be associated with different time zones with possible adjustments such as summer time. Time can be regular (such as quarterly economic data or daily weather data), or irregular (such as patient visits to a doctor's office). Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, and multiple grouping factors. Temporal data may comprise the occurrence of events, such as flight departures, that need to be reduced to a regular structure.

@@

@r4ds proposed the tidy data workflow, which provides a conceptual 
framework for processing data (as described in \autoref{fig:tsibble-fit}). Currently, time series modeling and forecasting enters this framework at the *modeling* stage, while temporal data enters at the start. This paper integrates
time series analysis into this tidy framework, providing a coherent way for getting temporal data into the matrix format for modeling.

```{r tsibble-fit, fig.cap = "(ref:tsibble-fit-cap)"}
include_graphics("img/tsibble-fit.pdf")
```

@@

The paper is structured as follows. \autoref{sec:str} reviews temporal data structures corresponding to time series and longitudinal analysis, and discusses "tidy data" and the grammar of data manipulation. \autoref{sec:semantics} proposes contextual semantics for temporal data, built on top of tidy data. The concept of data pipelines with respect to the time domain will be discussed in depth in \autoref{sec:pipeline}, followed by a discussion of the design choices made in the software structure in \autoref{sec:software}. Two case studies are presented in \autoref{sec:cases} illustrating temporal data exploration using the newly implemented infrastructure. \autoref{sec:conclusion} discusses future work.

@@

Current standards, provided by the native `ts` object in R, and extended by **zoo** [@zoo2005] and **xts** [@R-xts], assemble temporal data into matrices with implicit time indexes. These objects were designed for modeling methods. The diagram in the style of \autoref{fig:tsibble-fit} would place the model at the center of the analytical universe, and all the transformations and visualizations would hinge on that format. This is contrary to the **tidyverse** conceptualization, which holistically captures the full data workflow.

A new temporal data class is needed in the upstream of the workflow, which could incorporate all the downstream modules. A relatively new R package **tibbletime** [@R-tibbletime] proposed a data class of *time tibble* to represent temporal data in heterogeneous tabular format. It only requires an index variable to declare a temporal data object, thus placing it at the import stage. However, as proposed in \autoref{sec:semantics} a more rigid data structure is required for time series analytics and models.

@@

The choice of tidy representation of temporal data arises from a data-centric perspective, which accommodates all of the operations that are to be performed on the data. \autoref{fig:tsibble-fit} marks where this new abstraction is placed in the tidy model, which we refer to as a "tsibble". The tsibble structure is an extension of a data frame---a two-dimensional array in R---with additional time series semantics: index and key, as shown in \autoref{fig:tsibble-semantics}.

```{r tsibble-semantics, fig.cap = "(ref:tsibble-semantics-cap)"}
include_graphics("img/tsibble-semantics.pdf")
```

@@

To demonstrate the concept of the tsibble, \autoref{tab:tb-sub} presents a subset of tuberculosis cases estimated by @tb-data. It contains 12 observations and 5 variables arranged in a "long" tabular form. Each observation comprises the number of people who are diagnosed with tuberculosis for each gender at three selected countries in the years of 2011 and 2012. To turn this data into a tsibble: (1) column `year` is declared as the index variable; (2) the key is specified to consist of columns `country` and `gender`. The column `count` is the only measured variable in this data, but the structure is sufficiently flexible to hold other measured variables; for example, adding the corresponding population size (if known) in order to normalize the count later.

@@

The "key" is usually known a priori by analysts. For example, \autoref{tab:tb-sub} describes the number of tuberculosis cases for each gender across the countries every year. This data description suggests that columns `gender` and `country` have to be declared as the key, similar to a panel variable for longitudinal data. Lacking either of the two will be inadequate, because the observations would not be uniquely identified, and thus a tsibble construction would fail. An alternative specification of the key for this data is to include a third variable `continent`. Since `country` is nested within `continent`, it is a free variable for use. This variable brings additional information that can be used for forecasting reconciliation [@fpp]. The key needs to be explicit when multiple units exist in the data. The key can be implicit when it finds a univariate series in the table, but it cannot be absent from a tsibble.

The "key" also provides a link between the data, models, and forecasts. This neatly decouples the data from models and forecasts, leaving more room for necessary model components, such as coefficients, fitted values and residuals. More details are given  in \autoref{sec:ts-models}.

@@

\autoref{fig:tsibble-pipeline} illustrates the distinction of a time series pipeline from a regular data pipeline. It is highly recommended to check for identical entries of key and index before constructing a tsibble. Duplicates signal the data quality issue, which would likely affect subsequent analyses and hence decision making. Analysts are encouraged to gaze at data early and reason about the process of data cleaning. When the data meets the tsibble standard, it flows neatly into the analysis stage and takes full advantage of the tsibble infrastructure.

```{r tsibble-pipeline, fig.cap = "(ref:tsibble-pipeline-cap)"}
include_graphics("img/tsibble-pipeline.pdf")
```

@@

Besides the time gap verbs, the **tidyverse** vocabulary is adapted and expanded to facilitate time series transformations, as listed in \autoref{tab:verb-table}. The **tidyverse** suite showcases the general-purpose verbs for effectively manipulating tabular data, for example `filter()` picks observations, `select()` picks variables, and `left_join()` joins two tables. But these verbs need handling with care when used in the time series domain. A perceivable difference is summarizing variables between data frame and tsibble using `summarize()`. The former will reduce to a single summary, whereas the latter will obtain the index and their corresponding summaries. Users who are already familiar with the **tidyverse**, will experience a gentle learning curve for mastering these verbs and glide into time series analysis with low cognitive load.

@@

Modeling is crucial to explanatory and predictive analytics, but often imposes stricter assumptions on tsibble data. The verbs listed in \autoref{tab:verb-table} ease the transition to a tsibble that suits modeling. A tidy forecasting framework built on top of tsibble is under development, which aims at promoting transparent forecasting practices and concise model representation. A tsibble usually contains multiple time series. Batch forecasting will be enabled if a univariate model, such as ARIMA and Exponential Smoothing, is applied to each time series independently. This yields a "mable" (short for model table), where each model relates to each "key" value in tsibble. This avoids expensive data copying and reduces model storage. The mable is further supplied to forecasting methods, to produce a "fable" (short for forecasting table) in which each "key" along with its future time holds predictions. It also underlines the advantage of tsibble's "key" in acting as linkage between data inputs, models and forecasts. Advanced forecasting techniques, such as vector autocorrelation, hierarchical reconciliation, and ensembles, can be developed in a similar spirit. The modeling module is a current endeavor.

@@

Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data. The contextual summary provided by the print function, shown below on the data from \autoref{tab:tb-sub}, contains (1) data dimension with its shorthand time interval, alongside time zone if date-times, (2) variables that constitute the "key" with the number of series. These details aid users in understanding their data better.

@@

\autoref{fig:animate} shows the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia. A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series by computing average counts.

@@

Rolling window uses a programming paradigm---functional programming, which is different from those table verbs listed in \autoref{tab:verb-table}. Table verbs expect and return a tsibble, and does what the function name suggests. On the contrary, these rolling window functions could accept arbitrary input types and would return arbitrary sorts of output, depending on which method is put into the rolling window. For example, computing moving averages requires numerics and a function like `mean()`, and produces averaged numerics. However, rolling window regression takes a data frame and a linear regression method like `lm()`, and generates a complex object that contains coefficients, fitted values, and etc.

@@

Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution implemented in **nanotime** [@R-nanotime], and pure numbers in simulations. Ordered factors can also be a source of time, such as month names, January to December, and weekdays, Monday to Sunday. The **tsibble** package supports an extensive range of index types from numerics to nano time, but there might be custom indexes used for some occasions, for example school semesters. These academic terms vary from one institution to another, within an academic year which is defined differently from a calendar year. A new index would be immediately recognized by the software upon defining `index_valid()`, as long as it can be ordered from past to future. The interval regarding semesters is further outlined through `pull_interval()`. As a result, the rest of the software methods such as `has_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.

The class of tsibble is an underlying basis of temporal data, and there is a demand for sub-classing a tsibble. For example, a fable is actually an extension of a tsibble, mentioned in \autoref{sec:ts-models}. A low-level constructor `new_tsibble()` provides a vehicle to easily create a new subclass. This new object itself is a tsibble. It perhaps needs more metadata than those of a tsibble, that gives rise to a new data extension, like prediction distributions to a fable. Tsibble verbs are also S3 generics. Developers will be able to implement these verbs for the new class, if necessary.

@@

(ref:flight-pipeline1-cap) Flow chart illustrating the pipeline that preprocessed the data for creating \autoref{fig:carrier-mosaic}.

\autoref{fig:carrier-mosaic} visually presents the end product of aggregating the number of on-time and delayed flights to the year interval by carriers. This pipeline is initialized by defining a new variable if the flight is delayed, and involves summarizing the tallies of on-time and delayed flights for each carrier annually. To prepare the summarized data for a mosaic plot, it is further manipulated by melting new tallies into a single column. The flow chart (\autoref{fig:flight-pipeline1}) demonstrates the operations undertaking in the data pipeline. The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval (as each carrier ends up with only one annual summary). The final data set includes each carrier along with a single year, with the interval undetermined, which in turn feeds into the mosaic display. Note that Southwest Airlines (WN), as the largest carrier, operates less efficiently than Delta (DL).

@@

(ref:flight-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating \autoref{fig:sel-monthly-plot}.

What time of day and day of week should we travel to avoid suffering from horrible delay? \autoref{fig:draw-qtl} plots hourly quantile estimates across day of week in the form of small multiples. The upper-tail delay behaviors are of primary interest, and hence 50%, 80% and 95% quantiles are shown. To reduce the likelihood of suffering a delay, it is recommended to avoid the peak hour around 6pm (18). <!--As moving towards the upper extremes, the variations considerably increase, making departure time unpredictable.-->

@@

(ref:flight-pipeline3-cap) Flow chart illustrates the pipeline that preprocesses the data for creating \autoref{fig:draw-qtl}.

@@

During a power outage, electricity usage for some households may become unavailable, thus resulting in implicit missing values in the database. Gaps in time occur to `r scales::percent(sum(gap_df$.gaps) / NROW(gap_df))` of the households in this dataset. It would be interesting to explore these missing patterns as part of a preliminary analysis. Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths for each household. \autoref{fig:count-gaps} shows the gaps for the top 49 households arranged in rows from high to low in tallies. (The remaining households values have been aggregated into a single batch and appear at the top.) Missing values can be seen to occur at any time during the entire span. A small number of customers have undergone energy unavailability in consecutive hours, indicated by a line range in the plot. On the other hand, the majority suffer occasional outages with more frequent occurrence in January.

@@

Aggregation across all individuals helps to sketch a big picture of the behavioral change over time, organized into a calendar display (\autoref{fig:calendar-plot}). Each glyph represents the daily pattern of average residential electricity usage every thirty minutes. Higher consumption is indicated by higher values, and typically occurs in daylight hours. Color indicates hot days. The daily snapshots vary depending on the season in the year. During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially on hot days with daily average temperature greater than 25 degrees C. However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening. This plot illustrates how the tsibble data can easily integrate with other tools and graphics.

@@

(ref:elec-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating \autoref{fig:calendar-plot}.

@@

The authors would like to thank Mitchell O'Hara-Wild for many discussions on the software development and Davis Vaughan for contributing ideas on rolling window functions. We also thank Stuart Lee for the feedback on this manuscript. This article was created with knitr [@knitr] and R Markdown [@rmarkdown]. The project's Github repository <https://github.com/earowang/paper-tsibble> houses all materials required to reproduce this article and a history of the changes. <!-- To produce the analysis results as described in Section \autoref{sec:cases}, we require the **tsibble** package (`v0.6.2`) and the **dplyr** package (`v0.7.8`) installed. -->

>>>>

---
title: "A new tidy data structure to support exploration and modeling of temporal data"
author:
  - familyname: Wang
    othernames: Earo
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : earo.wang@monash.edu
    correspondingauthor: true
  - familyname: Cook
    othernames: Dianne
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : dicook@monash.edu
  - familyname: Hyndman
    othernames: Rob J
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : rob.hyndman@monash.edu
abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, and heterogeneous data types. On the other hand, the software supporting time series modeling and forecasting, makes strict assumptions on the data to be provided, typically requiring a matrix of numeric data with implicit time indexes. Going from raw data to model-ready data is painful. This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization, modeling and forecasting routines. Tidy data principles are extended to temporal data by: (1) mapping the semantics of a dataset into its physical layout; (2) including an explicitly declared index variable representing time; (3) incorporating a "key" comprising single or multiple variables to uniquely identify units over time. This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based contexts. A sound data pipeline facilitates a fluent workflow for analyzing temporal data. The infrastructure of tidy temporal data has been implemented in the R package **tsibble**.'
keywords: "time series, data wrangling, tidy data, R, forecasting, data science, exploratory data analysis, data pipelines"
wpnumber: no/yr
blind: false
cover: false
toc: false
bibliography: [references.bib]
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    dev: "pdf"
---

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold', 
  cache = FALSE, cache.path = 'cache/', 
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
knitr::opts_knit$set(root.dir = here::here())
```

```{r external, include = FALSE}
read_chunk('scripts/demo.R')
read_chunk('scripts/flights.R')
read_chunk('scripts/smart-meter.R')
```

@@

Temporal data arrives in many possible formats, with many different time contexts. For example, time can have various resolutions (hours, minutes, and seconds), and can be associated with different time zones with possible adjustments such as summer time. Time can be regular (such as quarterly economic data or daily weather data), or irregular (such as patient visits to a doctor's office). Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, and multiple grouping factors. Temporal data may comprise the occurrence of events, such as flight departures, that need to be reduced to a regular structure.

@@

@r4ds proposed the tidy data workflow, which provides a conceptual framework for processing data (as described in Figure \@ref(fig:tsibble-fit)). Currently, time series modeling and forecasting enters this framework at the *modeling* stage, while temporal data enters at the start. This paper integrates
time series analysis into this tidy framework, providing a coherent way for getting temporal data into the matrix format for modeling.

```{r tsibble-fit, fig.cap = "(ref:tsibble-fit-cap)"}
include_graphics("img/tsibble-fit.png")
```

@@

The paper is structured as follows. Section \@ref(sec:str) reviews temporal data structures corresponding to time series and longitudinal analysis, and discusses "tidy data" and the grammar of data manipulation. Section \@ref(sec:semantics) proposes contextual semantics for temporal data, built on top of tidy data. The concept of data pipelines with respect to the time domain will be discussed in depth in Section \@ref(sec:pipeline), followed by a discussion of the design choices made in the software structure in Section \@ref(sec:software). Two case studies are presented in Section \@ref(sec:cases) illustrating temporal data exploration using the newly implemented infrastructure. Section \@ref(sec:conclusion) discusses future work.

@@

Current standards, provided by the native `ts` object in R, and extended by **zoo** [@zoo2005] and **xts** [@R-xts], assemble temporal data into matrices with implicit time indexes. These objects were designed for modeling methods. The diagram in the style of Figure \@ref(fig:tsibble-fit) would place the model at the center of the analytical universe, and all the transformations and visualizations would hinge on that format. This is contrary to the **tidyverse** conceptualization, which holistically captures the full data workflow.

A new temporal data class is needed in the upstream of the workflow, which could incorporate all the downstream modules. A relatively new R package **tibbletime** [@R-tibbletime] proposed a data class of *time tibble* to represent temporal data in heterogeneous tabular format. It only requires an index variable to declare a temporal data object, thus placing it at the import stage. However, as proposed in Section \@ref(sec:semantics) a more rigid data structure is required for time series analytics and models.

@@

The choice of tidy representation of temporal data arises from a data-centric perspective, which accommodates all of the operations that are to be performed on the data. Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the tidy model, which we refer to as a "tsibble". The tsibble structure is an extension of a data frame---a two-dimensional array in R---with additional time series semantics: index and key, as shown in Figure \@ref(fig:tsibble-semantics).

```{r tsibble-semantics, fig.cap = "(ref:tsibble-semantics-cap)"}
include_graphics("img/tsibble-semantics.png")
```

@@

To demonstrate the concept of the tsibble, Table \@ref(tab:tb-sub) presents a subset of tuberculosis cases estimated by @tb-data. It contains 12 observations and 5 variables arranged in a "long" tabular form. Each observation comprises the number of people who are diagnosed with tuberculosis for each gender at three selected countries in the years of 2011 and 2012. To turn this data into a tsibble: (1) column `year` is declared as the index variable; (2) the key is specified to consist of columns `country` and `gender`. The column `count` is the only measured variable in this data, but the structure is sufficiently flexible to hold other measured variables; for example, adding the corresponding population size (if known) in order to normalize the count later.

@@

The "key" is usually known a priori by analysts. For example, Table \@ref(tab:tb-sub) describes the number of tuberculosis cases for each gender across the countries every year. This data description suggests that columns `gender` and `country` have to be declared as the key, similar to a panel variable for longitudinal data. Lacking either of the two will be inadequate, because the observations would not be uniquely identified, and thus a tsibble construction would fail. An alternative specification of the key for this data is to include a third variable `continent`. Since `country` is nested within `continent`, it is a free variable for use. This variable brings additional information that can be used for forecasting reconciliation [@fpp]. The key needs to be explicit when multiple units exist in the data. The key can be implicit when it finds a univariate series in the table, but it cannot be absent from a tsibble.

The "key" also provides a link between the data, models, and forecasts. This neatly decouples the data from models and forecasts, leaving more room for necessary model components, such as coefficients, fitted values and residuals. More details are given  in Section \@ref(sec:ts-models).

@@

Figure \@ref(fig:tsibble-pipeline) illustrates the distinction of a time series pipeline from a regular data pipeline. It is highly recommended to check for identical entries of key and index before constructing a tsibble. Duplicates signal the data quality issue, which would likely affect subsequent analyses and hence decision making. Analysts are encouraged to gaze at data early and reason about the process of data cleaning. When the data meets the tsibble standard, it flows neatly into the analysis stage and takes full advantage of the tsibble infrastructure.

```{r tsibble-pipeline, fig.cap = "(ref:tsibble-pipeline-cap)"}
include_graphics("img/tsibble-pipeline.png")
```

@@

Besides the time gap verbs, the **tidyverse** vocabulary is adapted and expanded to facilitate time series transformations, as listed in Table \@ref(tab:verb-table). The **tidyverse** suite showcases the general-purpose verbs for effectively manipulating tabular data, for example `filter()` picks observations, `select()` picks variables, and `left_join()` joins two tables. But these verbs need handling with care when used in the time series domain. A perceivable difference is summarizing variables between data frame and tsibble using `summarize()`. The former will reduce to a single summary, whereas the latter will obtain the index and their corresponding summaries. Users who are already familiar with the **tidyverse**, will experience a gentle learning curve for mastering these verbs and glide into time series analysis with low cognitive load.

@@

Modeling is crucial to explanatory and predictive analytics, but often imposes stricter assumptions on tsibble data. The verbs listed in Table \@ref(tab:verb-table) ease the transition to a tsibble that suits modeling. A tidy forecasting framework built on top of tsibble is under development, which aims at promoting transparent forecasting practices and concise model representation. A tsibble usually contains multiple time series. Batch forecasting will be enabled if a univariate model, such as ARIMA and Exponential Smoothing, is applied to each time series independently. This yields a "mable" (short for model table), where each model relates to each "key" value in tsibble. This avoids expensive data copying and reduces model storage. The mable is further supplied to forecasting methods, to produce a "fable" (short for forecasting table) in which each "key" along with its future time holds predictions. It also underlines the advantage of tsibble's "key" in acting as linkage between data inputs, models and forecasts. Advanced forecasting techniques, such as vector autocorrelation, hierarchical reconciliation, and ensembles, can be developed in a similar spirit. The modeling module is a current endeavor.

@@

Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data. The contextual summary provided by the print function, shown below on the data from Table \@ref(tab:tb-sub), contains (1) data dimension with its shorthand time interval, alongside time zone if date-times, (2) variables that constitute the "key" with the number of series. These details aid users in understanding their data better.

@@

Figure \@ref(fig:animate) shows the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia. A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series by computing average counts.

@@

Rolling window uses a programming paradigm---functional programming, which is different from those table verbs listed in Table \@ref(tab:verb-table). Table verbs expect and return a tsibble, and does what the function name suggests. On the contrary, these rolling window functions could accept arbitrary input types and would return arbitrary sorts of output, depending on which method is put into the rolling window. For example, computing moving averages requires numerics and a function like `mean()`, and produces averaged numerics. However, rolling window regression takes a data frame and a linear regression method like `lm()`, and generates a complex object that contains coefficients, fitted values, and etc.

@@

Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution implemented in **nanotime** [@R-nanotime], and pure numbers in simulations. Ordered factors can also be a source of time, such as month names, January to December, and weekdays, Monday to Sunday. The **tsibble** package supports an extensive range of index types from numerics to nano time, but there might be custom indexes used for some occasions, for example school semesters. These academic terms vary from one institution to another, within an academic year which is defined differently from a calendar year. A new index would be immediately recognized by the software upon defining `index_valid()`, as long as it can be ordered from past to future. The interval regarding semesters is further outlined through `interval_pull()`. As a result, the rest of the software methods such as `has_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.

The class of tsibble is an underlying basis of temporal data, and there is a demand for sub-classing a tsibble. For example, a fable is actually an extension of a tsibble, mentioned in Section \@ref(sec:ts-models). A low-level constructor `new_tsibble()` provides a vehicle to easily create a new subclass. This new object itself is a tsibble. It perhaps needs more metadata than those of a tsibble, that gives rise to a new data extension, like prediction distributions to a fable. Tsibble verbs are also S3 generics. Developers will be able to implement these verbs for the new class, if necessary.

@@

(ref:flight-pipeline1-cap) Flow chart illustrating the pipeline that preprocessed the data for creating Figure \@ref(fig:carrier-mosaic).

Figure \@ref(fig:carrier-mosaic) visually presents the end product of aggregating the number of on-time and delayed flights to the year interval by carriers. This pipeline is initialized by defining a new variable if the flight is delayed, and involves summarizing the tallies of on-time and delayed flights for each carrier annually. To prepare the summarized data for a mosaic plot, it is further manipulated by melting new tallies into a single column. The flow chart (Figure \@ref(fig:flight-pipeline1)) demonstrates the operations undertaking in the data pipeline. The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval (as each carrier ends up with only one annual summary). The final data set includes each carrier along with a single year, with the interval undetermined, which in turn feeds into the mosaic display. Note that Southwest Airlines (WN), as the largest carrier, operates less efficiently than Delta (DL).

@@

(ref:flight-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating Figure \@ref(fig:sel-monthly-plot).

What time of day and day of week should we travel to avoid suffering from horrible delay? Figure \@ref(fig:draw-qtl) plots hourly quantile estimates across day of week in the form of small multiples. The upper-tail delay behaviors are of primary interest, and hence 50%, 80% and 95% quantiles are shown. To reduce the likelihood of suffering a delay, it is recommended to avoid the peak hour around 6pm (18).

@@

(ref:flight-pipeline3-cap) Flow chart illustrates the pipeline that preprocesses the data for creating Figure \@ref(fig:draw-qtl).

@@

During a power outage, electricity usage for some households may become unavailable, thus resulting in implicit missing values in the database. Gaps in time occur to `r scales::percent(sum(gap_df$.gaps) / NROW(gap_df))` of the households in this dataset. It would be interesting to explore these missing patterns as part of a preliminary analysis. Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths for each household. Figure \@ref(fig:count-gaps) shows the gaps for the top 49 households arranged in rows from high to low in tallies. (The remaining households values have been aggregated into a single batch and appear at the top.) Missing values can be seen to occur at any time during the entire span. A small number of customers have undergone energy unavailability in consecutive hours, indicated by a line range in the plot. On the other hand, the majority suffer occasional outages with more frequent occurrence in January.

@@

Aggregation across all individuals helps to sketch a big picture of the behavioral change over time, organized into a calendar display (Figure \@ref(fig:calendar-plot)). Each glyph represents the daily pattern of average residential electricity usage every thirty minutes. Higher consumption is indicated by higher values, and typically occurs in daylight hours. Color indicates hot days. The daily snapshots vary depending on the season in the year. During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially on hot days with daily average temperature greater than 25 degrees C. However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening. This plot illustrates how the tsibble data can easily integrate with other tools and graphics.

@@

(ref:elec-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating Figure \@ref(fig:calendar-plot).

@@

The authors would like to thank Mitchell O'Hara-Wild for many discussions on the software development and Davis Vaughan for contributing ideas on rolling window functions. We also thank Stuart Lee for the feedback on this manuscript. This article was created with knitr [@knitr] and R Markdown [@rmarkdown]. The project's Github repository <https://github.com/earowang/paper-tsibble> houses all materials required to reproduce this article and a history of the changes.
